Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2024-10-25 Latex_files/Thesis/chapters/chapter 3/main.tex
updates

\chapter{Effects of localised shearing on crystal growth and nucleation}
As outlined in Chapter 1, the original outline of the project is to 
investigate the possibility of using optical tweezing to induce nucleation 
by generating fluid flow within a supersaturated solution. The intent 
of which would be twofold: Firstly to have a repeatable means of 
inducing nucleation under different solution conditions. And secondly, 
to understand the influence of shearing on nucleation at a micro 
level as compared to results in bulk fluid. It has been shown that 
for macro-scale systems, the likelihood of nucleation increases to 
a maximum value under increased shearing \cite{Debuysschere2023, 
Mura2016}. Mura and Zaccone developed a theoretical framework to 
describe how the a newly formed nucleus experiences two additional 
growth factors when placed in a moving fluid. Firstly, due to increased 
molecular transport of solute molecules the nucleation rate is enhanced 
in low to moderate fluid flows. But in addition, due to shear flow the 
crystal surface undergoes deformation which suppresses the nucleation 
rate undergoing faster fluid flow \cite{Mura2016}. The theory was 
confirmed for glycine solution by Debuysschere \textit{et al} who
identified that the nucleation rate of supersaturated glycine was 
enhanced up until $\dot{\gamma}\approx3000\ s^{-1}$ \cite{Debuysschere2023}.
After which the nucleation rate began to decrease but was 
still greater compared to the case where fluid flow was minimal.

Optical tweezers can been used to rotate a whole host of micro-rotors, 
with the fastest reported results exceeding $1000\ Hz$ in heavy water
\cite{Arita2016}. Therefore, it stands to reason that a micro-rotor
rotating in a supersaturated solution could generate sufficient fluid
flow such that the nucleation rate around said particle is enhanced 
compared to the bulk fluid. We focused on two primary candidates for
micro-rotors, Vaterite and 4-Heptyl-4-biphenylcarbonitrile (7CB). The
former being a polymorph of calcium carbonate and the latter an example
of nematic liquid crystals, both of which have been used repeatedly in
previous micro-rotor. In addition, we also consider the application of
using techniques beam steering to generate fluid flow by trapping silica
micro-beads. In this instance the fluid flow is generated not due to the
transfer of angular momentum, but due to shearing caused by a moving 
sphere through stagnant fluid.

To begin with, the discussion of the necessary optical equipment is 
covered, drawing attention to specialised components and techniques 
that are not standard in optical trapping set ups.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\section{Optical Tweezer Equipment}
In general, all optical tweezers require a laser driver, a focusing 
microscope objectives, a position controller, and position detector.
The laser used for this project was a $1064\ nm$ near infrared laser 
- provided by CNI Lasers – that was focused by a Nikon 100x oil 
immersion lens. The choice of an oil immersion lens is important as 
the optical oil used prevents a loss of focus when used on a glass 
cover slip. Now, experimental work has shown that the trapping 
efficiency increases with beam diameter up until it exceeds 
$\frac{2}{3}D_{obj}$ \cite{kim2003dependence} where $D_{obj}$ is 
the diameter of the objective aperture. To expand the beam front we 
utilise a Galilean beam expansion arrangement (indicated by $f_1$, 
and $f_2$ in Fig.~\ref{fig:setup}) as recommended for high power laser 
applications. In our initial experiments the beam expansion provides a 
$4\times$ magnification. Whereas in later experiments we utilised a 
galvano-mirror the beam expansion is $3\times$ and then the 4f 
correlator provides a further $1.25\times$ magnification (using $f_3$ 
and $f_4$) - the magnification is given by.
\begin{align}
	\frac{D_2}{D_1} = \frac{f_2}{f_1}
\end{align}

It should be noted that the galvano-mirror requires the use of a 
Keplerian beam expansion arrangement which reduces the transmitted 
laser power due to localised heating of the air. Afterwards the 
laser is passed through a dichroic mirror that separates incoming 
infrared and visible light, this is to prevent the laser from 
damaging the CCD camera used for imaging the trapping plane. The 
laser is then focused to a diffraction limited spot by the objective. 
Utilizing a high numerical aperture objective enhances the gradient 
force at the focal point; the trade-off being that the for higher NA 
objectives the trapping depth is reduced due to spherical aberrations. 
While it is possible to increase the trapping depth \cite{Reihani2006} 
by adjusting the objective's tube length this approach is incompatible 
with our trapping arrangement. A 0.25 NA condenser objective refocuses 
the scattered laser light and also provide an aperture for an imaging 
LED to illuminate the focal plane. Samples are loaded onto a piezo 
driven table to that is inserted between the trapping and condensing 
objectives; the piezo drivers allow for sub-micron control of the 
beam focus position to a degree as small as a 10 nm. To detect and 
monitor the position of a trapped particle a quadrant photo diode 
(QPD) was utilised. 
\begin{figure}[h!]
	\centering
	\includegraphics[height=\linewidth, angle=270]{tweezer_setup.pdf}
	\caption{Optical tweezer set up used for the majority of the PhD. The focal lengths of $f_1$, $f_2$, $f_3$, \& $f_4$ are $-20\ mm$, $60\ mm$, $150\ mm$, \& $200\ mm$ respectively. Diagram not drawn to scale.}
	\label{fig:setup}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Position detection methods}
In order to accurately capture the dynamics of a trapped particle, a 
position detection system is required. There are 3 possible methods 
of position detection: video-analysis, lateral-effect position sensing, 
and photodiodes. The former being ideally suited for multiple traps 
or situations where precision is not the top priority. In order to 
match the force measurements of back-focal plane interferometry requires 
the camera's frame rate to exceed $1\ kHz$ which can be difficult 
to achieve while maintaining a decent resolution \cite{Gibson2008}. 
In comparison off the shelf back-focal plane detectors can achieve 
temporal resolutions anywhere from $10-100\ kHz$ \cite{BergSoerensen2004}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fourier Optics and 4f correlators}
A 4f correlator is an example of Fourier optics in practice, understanding
that a focused lens takes a Fourier transform of the light profile. 
Consider a laser with a circular Gaussian profile, if you were to place 
a detector there you would pick up the intensity as a function of its 
position within the beam. If however you focused the light into a single 
point (using a +ve focal lens) you are actually seeing a measurement of 
the phase of your laser with position, in which you would see a diffraction
limited spot ($d = \lambda/2nsin(\theta)$), indicating that the laser is 
collimated. In imaging systems, a series of focal lenses can be used to 
filter out unwanted scattering from an image (or in an inverse case 
differentiate between different images), the placement of each lens is 
shown below.
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.475\linewidth}
		\includegraphics[width=\linewidth]{signal_generator.jpg}
		\caption{}
	\end{subfigure}
	\begin{subfigure}{0.475\linewidth}
		\includegraphics[width=\linewidth, height=0.7\linewidth]
		{galvano_mirror_controllers.jpg}
		\caption{}
	\end{subfigure}

	\caption{Signal generator galvano mirror controller, channel 1 controls the x-axis 
		mirror, while channel 2 controls the y-axis mirror. Both channels can be manipulated
		independently.}
\end{figure}

For our applications a 4f correlator is utilised to ensure that the motion
of the galvano-mirrors does not move the focal point of the laser, allowing 
for a stable trap even while in motion. As shown in Fig.~\ref{fig:setup}, after
the galvano-mirror we have our two lenses - $f_3$ and $f_4$ - the former being 
installed $150\ mm$ from the second mirror of the galvano, and the latter being 
installed $200\ mm$ from the back focal plane of the trapping objective. The 
signal generator used was supplied by 'MCS Test Equipment Ltd', allowing for 
dual channel signal control. This allowed us to precisely control the alignment,
amplitude, phase, and frequency of both mirrors making alignment much easier. For
basic trapping calibration the galvano-mirrors were set to a simple DC output, 
providing a fixed spot which operates like a typical optical trap.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Synthesis of Birefringent Micro spheres}
\label{sec:vaterite}
There are several options for particles that can be rotated using optical 
tweezers \cite{Parkin2009, Saito2022}. Over the course of the project two 
different micro spheres where investigated, Vaterite and liquid crystal 
droplets. Both can be readily synthesised in the lab and are will rotate 
at a variety of sizes.

Vaterite is a polymorph of calcium carbonate that is rarely seen in nature 
due to its low stability \cite{KonopackaLyskawa2019}. However unlike its 
other polymorphs of calcite and aragonite, when synthesised Vaterite will 
typically form small spherical particles making them ideal for optical 
trapping and rotation. Synthesis of Vaterite micro spheres requires fine 
control of the crystal growth process in order to maintain polymorphic stability. 
Though for the purposes of optical rotation the exact polymorph is not as 
important as its morphology as all 3 polymorphs are inherently birefringent.

Vaterite samples where made by the first preparing equal amounts of 
$CaCl_2$ and $Na_2CO_3$ at a concentration of $0.33M$, at the same time
a vial of $0.33M\ MgSO_4$ was prepared and set aside for later. First
a small vial was filled with $1.5mL$ of $CaCl_2$ followed by $60\mu L$ 
and $90\mu L$ of $MgSO_4$ and $NaCO_3$ respectively, forming a seed solution. 
Next, a larger vial was filled with $5\ mL$, $1.5\ mL$, and $1\ mL$ of
$CaCL_2$, $MgSO_4$, and $NaCO_3$ respectively followed by the seed solution. 
After 10 minutes of slow but continuous mixing a few drops of Agepon was added
to halt the reaction, the solution was filtered and washed 3 times with 
distilled water before being suspended in water.

When trapped in circularly polarised light, the anisotropic crystal lattice
allows spin angular momentum to be transferred to the Vaterite particle, 
resulting in a rotation about the beam axis. In addition, the anisotropic 
scattering causes the QPD signal to vary with a constant periodicity that
is attributed to its rotation rate. Therefore, the resulting power spectrum 
is not a Lorentzian but now also displays peaks that appear at integer 
multiples of the particles rotational frequency. 
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.4\linewidth}
		\includegraphics[width=\linewidth]{vaterite_sample.jpg}
		\subcaption{}
	\end{subfigure}
	\begin{subfigure}{0.55\linewidth}
		\includegraphics[width=\linewidth]{rotating_psd.png}
		\subcaption{}
	\end{subfigure}
	\caption{(a)Sample Vaterite sphere suspended in water and trapped by circular polarised trap. (b) Collected power spectrum from rotating Vaterite, peaks in the power spectrum appear at integer multiples of the rotational frequency ($f_{rot} \approx 49.8$ Hz)}
	\label{fig:vaterite}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Rotation of birefringent micro spheres} 
Optical tweezing has often been used for micro-rheology, by computing the
exact forces being exerted on the trapped sphere, one can determine the
local temperature/viscosity of the medium \cite{Millen2014, RodriguezSevilla2018}.
Using a birefringent particle and rotating it within the fluid, the maximum 
rotation rate is due to the fluid drag resisting the torque of the trapping 
beam \cite{RodriguezSevilla2018}. If you want to measure fluid flow you can 
instead use a micro-rotor to see how fluid flow propagates in the medium \cite{Knoener2005}. Likewise, one can use a beam steering arrangement to 
probe the drag force of the fluid, by understanding the trap strength 
(calibrating using a low frequency signal) one can measure the drag force 
experienced by the local fluid \cite{RobertsonAnderson2018}. I

Understanding the fluid velocity around our trapped object is determined 
mostly by the Reynold's number of the system, for a sphere submersed in a 
moving fluid of velocity $U$ this is given by:
\begin{align}
	Re = \frac{\rho UD}{\mu}
\end{align}

Where $D$ is the sphere's diameter, and $\rho$ and $\mu$ are the fluid's 
density and viscosity respectively. In our case we do not have a fluid
moving around a sphere but a sphere moving through the fluid at some 
velocity $U$, assuming a no-slip boundary condition we can model the 
fluid velocity profile based on the velocity of the particle. There are 
two possible avenues for generating shear flow with a trapped particle; 
rotation of birefringent particles, and fluid flow induced by particle motion.

Rotating birefringent particles are by far the most common method for 
generating and measuring fluid flow in a solution. To see if we can 
even achieve the theoretical maximum shear rate, Vaterite spheres 
were synthesised (see Sec.\ref{sec:vaterite}) submerged in water and 
trapped with the 1064 nm laser at set to 450 mW. The rotation frequency 
was determined using the QPD, and the particle sizes were computed 
by image analysis. With the particle size and rotation frequency, the 
tangential rotation speed is calculated via:
\begin{align}
	\label{eq:birefringent_speed}
	u(r) = \frac{\pi}{4}\frac{d^3}{r^2}\omega
\end{align}

Where $d$ is the particle diameter, $\omega$ is the rotation frequency
reported by the QPD, and $r$ is the distance from the particle's centre. 
Using Eq.\ref{eq:birefringent_speed} we calculated the fluid flow radiating
outward from the centre of the sphere. The shear rate can then be computed
as the partial derivative fluid flow (assuming shearing is generated purely
by the flow field):
\begin{align}
	\label{eq:birefringent_shear}
	\dot{\gamma}(r)=\left|\frac{\delta u(r)}{\delta r} \right|= \frac{\pi}{2}\frac{d^3}{r^3}\omega
\end{align}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Estimation of fluid flow around micro-rotors in bulk fluid}
First we determined the upper rotation rate that could be achieved 
using both Vaterite and liquid crystal spheres. Vaterite samples were 
synthesised according to \cite{Parkin2009, Bishop2004} (see sec.~
\ref{sec:vaterite}), and then suspended in distilled water. A sample
of $200\ \mu L$ was pipetted and a single microsphere was captured via
a circular polarised beam.

Due to Van der Waal's forces some of the microspheres were stuck together, 
fortunately individual sphere's were still present. Multiple microspheres 
were trapped and their rotation rate was determined by looking at the 
peak frequency component of the collected power spectrum. The shear flow
was estimated using eq.\eqref{eq:birefringent_shear} assuming the spheres
were operating in the bulk fluid and away from any boundaries.
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.75\linewidth}
		\includegraphics[width=\linewidth]{vaterite_fluid_flow.png}
	\end{subfigure}
	\begin{subfigure}{0.75\linewidth}
		\includegraphics[width=\linewidth]{vaterite_shear_rate.png}
	\end{subfigure}
	\caption{(Top) Fluid flow radiating out from the surface of a rotating 
		Vaterite sphere. (Bottom) Shear rates computed using Eq.\ref{eq:birefringent_shear}, optimal shear rate is of $3000 s^{-1}$ 
		is indicated by the dotted line. Vaterite radii and rotation frequencies 
		are shown, the laser power was kept constant at 450 mW. Reported rotation
		rates, and their corresponding fluid flow and shear rates, for Vaterite 
		are also plotted alongside lab results. Results from \cite{Parkin2009, Arita2016} are included as well.}
	\label{fig:vaterite_shear}
\end{figure}

>>>>

\chapter{Effects of localised shearing on crystal growth and nucleation}
As outlined in Chapter 1, part of the projects aim is to investigate 
the possibility of using optical tweezing to induce nucleation by 
generating fluid flow within a supersaturated solution. The intent 
of which would be twofold: Firstly to have a repeatable means of 
inducing nucleation under different solution conditions. And secondly, 
to understand the influence of shearing on nucleation at a micro 
level as compared to results in bulk fluid.

It has been shown that for macro-scale systems, the likelihood of 
nucleation increases to a maximum value under increased shearing 
\cite{Debuysschere2023, Mura2016}. Mura and Zaccone developed a 
theoretical framework to describe how the a newly formed nucleus 
experiences two additional growth factors when placed in a moving 
fluid. Firstly, due to increased molecular transport of solute 
molecules the nucleation rate is enhanced in low to moderate fluid 
flows. But in addition, due to shear flow the crystal surface 
undergoes deformation which suppresses the nucleation rate 
undergoing faster fluid flow \cite{Mura2016}. Experimental results 
with glycine solution support this theory; Debuysschere \textit{et 
al} demonstrated that the nucleation rate of supersaturated glycine 
was enhanced up until $\dot{\gamma}\approx3000\ s^{-1}$ 
\cite{Debuysschere2023}. After which the nucleation rate began to 
decrease but was still greater compared to the case where fluid 
flow was minimal.

Optical tweezers can been used to rotate a whole host of microscopic 
objects, with the fastest reported results exceeding $1000\ Hz$ in 
heavy water \cite{Arita2016}. If such a micro-rotor was suspended in
a supersaturated solution the fluid flow around it could be fast enough
that the nucleation rate is locally enhanced. We focused on two 
primary candidates for rotation, Vaterite and 
4-Heptyl-4-biphenylcarbonitrile (7CB). The former being a polymorph of 
calcium carbonate and the latter an example of nematic liquid crystals, 
both of which have been used repeatedly in previous micro-rotor research
\cite{Saito2022, Arita2016, Parkin2009}. In addition, we also consider 
the application of using techniques beam steering to generate fluid 
flow by trapping silica micro-beads. In this instance the fluid flow is 
generated not due to the transfer of angular momentum, but due to 
shearing caused by a moving sphere through stagnant fluid.

To begin with, the discussion of the necessary optical equipment is 
covered, drawing attention to specialised components and techniques 
that are not standard in optical trapping set ups.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\section{Optical Tweezer Equipment}
In general, all optical tweezers require a laser driver, a focusing 
microscope objectives, a position controller, and position detector 
\cite{Gieseler2020}. The laser used for this project was a $1064\ 
nm$ near infrared laser - provided by CNI Lasers – that was focused 
by a Nikon 100x oil immersion lens. The choice of an oil immersion 
lens is important as the optical oil used prevents a loss of focus 
when used on a glass cover slip. Now, experimental work has shown 
that the trapping efficiency increases with beam diameter up until 
it exceeds $\frac{2}{3}D_{obj}$ \cite{kim2003dependence} where 
$D_{obj}$ is the diameter of the objective aperture. To expand the 
beam front we utilise a Galilean beam expansion arrangement 
(indicated by $f_1$, and $f_2$ in Fig.~\ref{fig:setup}) as recommended 
for high power laser applications. In our initial experiments the 
beam expansion provides a $4\times$ magnification. Whereas in later 
experiments we utilised a galvano-mirror the beam expansion is 
$3\times$ and then the 4f correlator (see \ref{sec:4f_correlation})
provides a further $1.25\times$ magnification (using $f_3$ and $f_4$) 
- the magnification is given by.
\begin{align}
	\frac{D_2}{D_1} = \frac{f_2}{f_1}
\end{align}

It should be noted that the galvano-mirror requires the use of a 
Keplerian beam expansion arrangement which reduces the transmitted 
laser power due to localised heating of the air. Afterwards the 
laser is passed through a dichroic mirror that separates incoming 
infrared and visible light, this is to prevent the laser from 
damaging the CCD camera used for imaging the trapping plane. The 
laser is then focused to a diffraction limited spot by the objective. 
Utilizing a high numerical aperture objective enhances the gradient 
force at the focal point; the trade-off being that the for higher NA 
objectives the trapping depth is reduced due to spherical aberrations. 
While it is possible to increase the trapping depth \cite{Reihani2006} 
by adjusting the objective's tube length this approach is incompatible 
with our trapping arrangement. A 0.25 NA condenser objective refocuses 
the scattered laser light and also provide an aperture for an imaging 
LED to illuminate the focal plane. Samples are loaded onto a piezo 
driven table to that is inserted between the trapping and condensing 
objectives; the piezo drivers allow for sub-micron control of the 
beam focus position to a degree as small as a 10 nm. To detect and 
monitor the position of a trapped particle a quadrant photo diode 
(QPD) was utilised. 
\begin{figure}[h!]
	\centering
	\includegraphics[height=\linewidth, angle=270]{tweezer_setup.pdf}
	\caption{Optical tweezer set up used for the majority of the PhD. 
		The focal lengths of $f_1$, $f_2$, $f_3$, \& $f_4$ are $-20\ mm$, 
		$60\ mm$, $150\ mm$, \& $200\ mm$ respectively [M = mirrors,
		DM = dichroic mirrors, f = focal lenses]. Diagram not 
		drawn to scale.}
	\label{fig:setup}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Position detection methods}
In order to accurately capture the dynamics of a trapped particle, a 
position detection system is required. There are 3 possible methods 
of position detection: video-analysis, lateral-effect position sensing, 
and photodiodes.

Video analysis is ideally suited for multiple traps or situations where 
precision is not the top priority. Whereas lateral-effect and photodiode 
position detection are two examples of back-focal plane interferometry, 
where the interference pattern produced by the target particle is 
extrapolated to determine its position. In order for video analysis to 
match the accuracy of back-focal plane interferometry methods requires 
the camera's frame rate to exceed $1\ kHz$ which can be difficult to 
achieve while maintaining a decent resolution \cite{Gibson2008}. In 
comparison off the shelf back-focal plane detectors can achieve temporal 
resolutions anywhere from $10-100\ kHz$ \cite{BergSoerensen2004}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fourier Optics and 4f correlators}
\label{sec:4f_correlation}
As shown in \ref{fig:setup}, after the Galvano mirror there are a pair
of focal lenses ($f_3$ and $f_4$) that do not seem to serve a clear
purpose. These are in fact crucial for the operation of the Galvano, the
pair of them can be more accurately called a 4f correlator.

A 4f correlator is an example of Fourier optics in practice, understanding
that a focused lens takes a Fourier transform of the light profile. 
Consider a laser with a circular Gaussian profile, if you were to place 
a detector there you would pick up the intensity as a function of its 
position within the beam. If however you focused the light into a single 
point (using a +ve focal lens) you are actually seeing a measurement of 
the phase of your laser with position, in which you would see a diffraction
limited spot ($d = \lambda/2nsin(\theta)$), indicating that the laser is 
collimated. In imaging systems, a series of focal lenses can be used to 
filter out unwanted scattering from an image (or in an inverse case 
differentiate between different images), the placement of each lens is 
shown via fig.~\ref{fig:signal_generator}.
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.475\linewidth}
		\includegraphics[width=\linewidth]{signal_generator.jpg}
		\caption{}
	\end{subfigure}
	\begin{subfigure}{0.475\linewidth}
		\includegraphics[width=\linewidth, height=0.7\linewidth]
		{galvano_mirror_controllers.jpg}
		\caption{}
	\end{subfigure}

	\caption{Signal generator galvano mirror controller, channel 1 controls the x-axis 
		mirror, while channel 2 controls the y-axis mirror. Both channels can be manipulated
		independently.}
	\label{fig:signal_generator}
\end{figure}

For our applications a 4f correlator is utilised to ensure that the motion
of the galvano-mirrors does not move the focal point of the laser, allowing 
for a stable trap even while in motion. As shown in Fig.~\ref{fig:setup}, after
the galvano-mirror we have our two lenses - $f_3$ and $f_4$ - the former being 
installed $150\ mm$ from the second mirror of the galvano, and the latter being 
installed $200\ mm$ from the back focal plane of the trapping objective. The 
signal generator used was supplied by 'MCS Test Equipment Ltd', allowing for 
dual channel signal control. This allowed us to precisely control the alignment,
amplitude, phase, and frequency of both mirrors making alignment much easier.

While in theory the frequency can be increased until the mechanical limit of the 
mirrors is reached ($\approx 1000\ Hz$), the practical upper limit is determined 
by the trap strength. For a silica sphere this is on the order of $300\ Hz$ when
suspended in water. Likewise while in theory the signal amplitude can be increased
until the voltage limit of the motors is reached ($\approx 5.0 V$) the geometry 
of the focal lenses limits the maximum amplitude to $0.5 V$ as any greater will
move the laser beyond the lens. For basic trapping calibration the galvano-mirrors 
were set to DC output, providing a fixed spot which operates like a typical optical 
trap.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Synthesis of Birefringent Micro spheres}
\label{sec:vaterite}
There are several options for particles that can be rotated 
using optical tweezers \cite{Parkin2009, Saito2022}. Over 
the course of the project two different micro spheres where 
investigated, Vaterite and liquid crystal droplets. Both can 
be readily synthesised in the lab and are will rotate at a 
variety of sizes (see \ref{sec:opt_torque}).

Vaterite is a polymorph of calcium carbonate that is rarely 
seen in nature due to its low stability \cite{KonopackaLyskawa2019}. 
All three polymorphs are inherently birefringent meaning 
that they can be rotated using circularly polarised light. 
However unlike its other polymorphs of calcite and aragonite, 
when synthesised vaterite will typically form small spherical 
particles making them ideal for optical trapping and rotation. 
Synthesis of Vaterite micro spheres requires fine control of 
the crystal growth process in order to maintain polymorphic 
stability. Though for the purposes of optical rotation the 
exact polymorph is not as important as its morphology as all 
3 polymorphs are inherently birefringent.

Vaterite samples where made by the first preparing equal amounts 
of $CaCl_2$ and $Na_2CO_3$ at a concentration of $0.33M$, at 
the same time a vial of $0.33M\ MgSO_4$ was prepared and set 
aside for later. Firsta small vial was filled with $1.5mL$ of 
$CaCl_2$ followed by $60\mu L$ and $90\mu L$ of $MgSO_4$ and 
$NaCO_3$ respectively, forming a seed solution. Next, a larger 
vial was filled with $5\ mL$, $1.5\ mL$, and $1\ mL$ of $CaCL_2$, 
$MgSO_4$, and $NaCO_3$ respectively followed by the seed solution. 
After 10 minutes of slow but continuous mixing a few drops of 
Agepon was added to halt the reaction, the solution was filtered 
and washed 3 times with distilled water before being suspended 
in water.

When trapped in circularly polarised light, the anisotropic 
crystal lattice allows spin angular momentum to be transferred 
to the Vaterite particle, resulting in a rotation about the 
beam axis. Because the particle scatters light anisotropically 
the signal detected by the QPD shows periodic fluctuations 
\cite{Yogesha2012}. In addition due to the change in polarisation 
there is also a periodic variation that appears at twice the 
rotational frequency \cite{Monteiro2018}. Therefore, the resulting 
power spectrum is not a Lorentzian but now also displays peaks 
that appear at integer multiples of the particles rotational 
frequency. 
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.4\linewidth}
		\includegraphics[width=\linewidth]{vaterite_sample.jpg}
		\subcaption{}
	\end{subfigure}
	\begin{subfigure}{0.55\linewidth}
		\includegraphics[width=\linewidth]{rotating_psd.png}
		\subcaption{}
	\end{subfigure}
	\caption{(a)Sample Vaterite sphere suspended in water and trapped by
		 circular polarised trap. (b) Collected power spectrum from 
		 rotating Vaterite, peaks in the power spectrum appear at integer 
		 multiples of the rotational frequency ($f_{rot} \approx 49.8$ Hz)}
	\label{fig:vaterite}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Rotation of birefringent micro spheres} 
Optical tweezing has often been used for micro-rheology, by computing the
exact forces being exerted on the trapped sphere, one can determine the
local temperature/viscosity of the medium \cite{Millen2014, RodriguezSevilla2018}.
When it comes to optical rotation, a rotating particle will experience a
fluid drag torque that is proportional to its rotation rate, the maximum 
rotation rate is when the fluid drag is equal to the optical torque 
of the laser \cite{RodriguezSevilla2018}. Optical rotors have been used
to measure local temperature changes \cite{Millen2014}, or for understanding 
how fluid shear propagates through a medium \cite{Knoener2005}. Likewise, 
one can use a beam steering arrangement to probe the drag force of the 
fluid, by understanding the trap strength (calibrating using a low 
frequency signal) one can measure the drag force experienced by the local 
fluid \cite{RobertsonAnderson2018}. I

For any discussion of fluid flow its important that we know the fluid 
regime which is given by the Reynolds number. For a sphere submersed in a 
moving fluid of velocity $U$ this is given by:
\begin{align}
	Re = \frac{\rho UD}{\mu}
\end{align}

Where $D$ is the sphere's diameter, and $\rho$ and $\mu$ are the fluid's 
density and viscosity respectively. In our case we do not have a fluid
moving around a sphere but a sphere moving through the fluid at some 
velocity $U$, assuming a no-slip boundary condition we can model the 
fluid velocity profile based on the velocity of the particle. Assuming
that for now the fluid properties are unaffected by the rotating particle
we see that for an individual sphere the Reynolds number can only change 
due to increasing/decreasing the fluid velocity. The fluid velocity is 
directly proportional to the particles rotation rate for low Reynolds
numbers \cite{Bruce2020}. Given the small size of the particles used 
($1-10\mu\ m$ in diameter), and the relatively low rotation rates that 
are possible with an optical tweezer, for the case of our analysis the
Reynolds number will not never be large enough to consider factors such
as turbulence. There are two possible avenues for generating shear flow 
with a trapped particle; rotation of birefringent particles, and fluid 
flow induced by particle motion.

Rotating birefringent particles are the more common method for generating 
and measuring fluid flow in a solution. To see if we can even achieve the 
theoretical maximum shear rate, Vaterite spheres were synthesised (see 
Sec.\ref{sec:vaterite}) submerged in water and trapped with the 1064 nm 
laser at set to 450 mW. The rotation frequency was determined using the 
QPD, and the particle sizes were computed by image analysis. With the 
particle size and rotation frequency, the tangential rotation speed is 
calculated via:
\begin{align}
	\label{eq:birefringent_speed}
	u(r) = \frac{\pi}{4}\frac{d^3}{r^2}\omega
\end{align}

Where $d$ is the particle diameter, $\omega$ is the rotation frequency
reported by the QPD, and $r$ is the distance from the particle's centre. 
Using Eq.\ref{eq:birefringent_speed} we calculated the fluid flow radiating
outward from the centre of the sphere. The shear rate can then be computed
as the partial derivative fluid flow (assuming shearing is generated purely
by the flow field). This approach was used previously with liquid crystal 
rotors \cite{Saito2022}:
\begin{align}
	\label{eq:birefringent_shear}
	\dot{\gamma}(r)=\left|\frac{\delta u(r)}{\delta r} \right|= 
	\frac{\pi}{2}\frac{d^3}{r^3}\omega
\end{align}

One would expect that we set some reference distance where the fluid flow 
would be considered zero. However, at a microscopic level we are more 
concerned by the fluid shear for an individual nucleation event. We are
essentially computing what the shear rate would be for a small volume 
of fluid between two moving parallel plates. This is similar logic to Mura 
and Zaccone's work where they considered the shear on a subcritical nucleus
in a flow field \cite{Mura2016}. They go further in depth to discuss the 
impact of material transport, overall coming to the conclusion that 
increasing shear rate leads to an increase the nucleation rate up till some
maximum. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Estimation of fluid flow around a rotating particle in bulk fluid}
First we determined the upper rotation rate that could be achieved 
using both Vaterite and liquid crystal spheres. Vaterite samples 
were synthesised according to \cite{Parkin2009, Bishop2004} (see 
sec.~\ref{sec:vaterite}), and then suspended in distilled water. 
A sample of $200\ \mu L$ was pipetted and a single microsphere was 
captured via a circular polarised beam.

Due to Van der Waal's forces some of the microspheres were stuck 
together, fortunately individual sphere's were still present. 
Individual microspheres were trapped and their rotation rate was 
determined by looking at the peak frequency component of the 
collected power spectrum. Due to their anisotropic scattering the 
QPD signal will oscillate periodically. After taking the Fourier
transform of the QPD signal the power spectrum has distinctive
peaks at integer multiples of the rotational frequency. The largest 
peak is often twice the rotational frequency which is due to the 
polarization modulation of the scattered field, the first peak 
is due purely to the rotation of vaterite sphere \cite{Monteiro2018}. 
An example of this is behaviour is shown in fig.~\ref{fig:tdat_fdat}.

\begin{figure}[h!]
	\begin{subfigure}{0.49\linewidth}
		\includegraphics[width=\linewidth]{time_domain_data.png}
	\end{subfigure}
	\begin{subfigure}{0.49\linewidth}
		\includegraphics[width=\linewidth]{frequency_domain_data.png}
	\end{subfigure}
	\caption{Example of QPD signal being converted from the time 
		domain to the frequency domain. (left) Raw signal collected
		from the QPD over the 1st second of its trajectory showing 
		a periodic spacing with a period of $1/f_{rot}$. (right) 
		Power spectrum of the QPD signal, the y-axis is a linear 
		scale to demonstrate the clear peaks in the power spectra. 
		The rotational frequency is the first instance of a peak 
		forming, and after that the peaks appear in integer 
		multiples of $f_{rot}$}
	\label{fig:tdat_fdat}
\end{figure}
